---
title: "Final Project Linear Models"
output: html_document
---

```{r Load Packages and Data}
#Load packages
library(dplyr)
library(gridExtra)
library(corrplot)
library(ggplot2)
library(ggfortify)
#Load data
cuyama<-read.csv("Data/Cuyama.csv")

```

# Question 1

## Do latitude, longitude, shrub volume, and treatment affect microsite temperature and soil moisture? 

```{r Q1 - Data Wrangling}

#Filter data - keep columns shrub ID, x, y, z, treatment, measure,sensor, site and microsite. Site corresponds to latitude and longitude of the shrub. In contrast to our proposal, we will consider site to be categorical, as there are 6 distinct sites in our data.

cuyamafilt<-select(cuyama,site,shrub.ID,x,y,z,measure,treatment,sensor,microsite)

#Get mean temperature/soil moisture for each shrub in each microsite
cuyamameans<-summarize(group_by(cuyamafilt,shrub.ID,sensor,treatment,microsite), MeanMeasure=mean(measure,na.rm=TRUE))

#Get treatment, site, x, y and z for each shrub in each microsite
shrubstats<-distinct(cuyamafilt,shrub.ID,sensor,treatment,site,x,y,z,microsite)

#Get shrub volumes
shrubstats<-mutate(shrubstats, Volume=(2/3)*pi*(x/2)*(y/2)*z)

#Combine dataframe with mean temps/soil moistures with the shrub stats. No longer need x, y or z as we are using volume.
shrubstats<-cbind(shrubstats[order(shrubstats$shrub.ID,shrubstats$sensor,shrubstats$microsite),],MeanMeasure=cuyamameans$MeanMeasure)
shrubstats<-shrubstats[,-c(5:7)]
```

No missing data once filtered, do see some outliers for measures of shrub 164 (one soil moisture measure reading looks like a temperature reading and temperature looks like soil moisture) by a quick look at the data.

#### Check for bias in sample sizes among predictors site, treatment and microsite (volume should differ for each shrub so no bias). Check for correlations among numerical predictors.

```{r Q1 - Predictors}
samplesizesitefunc<-function(i){
  sum(shrubstats$site==i)}
samplesizessite<-sapply(1:6,samplesizesitefunc) #Sample sizes for site are fairly even except the first location has only two data points. 2 data points are a very low sample size, so we will need to look out for this in our analyses.

sum(shrubstats$microsite=="open") #As expected, the number of open microsites is about half of the data points (36/70 rows).
sum(shrubstats$treatment=="clipped") #sample size of clipped is 42 of 70 total rows, so we have a fairly even spread of clipped and unclipped data.

#We conclude that predictors do not show much bias, but will be wary of site 1.

#Check our only numerical predictor is shrub volume so no need to check for correlations. We can check how volume changes with site, just to be sure:

t<-as.data.frame(cor(shrubstats[,c(4,6)]))

#Use tableGrob from gridExtra package to create our table and store it as an object called "table"
table<-tableGrob(t, rows = rownames(t), cols = colnames(t), theme = ttheme_default(rowhead=list(
        fg_params=list(fontface="bold"),
        bg_params=list(fill=c(NA,rep("grey80",4)))
    )))
grid.arrange(table)

corrplot(cor(shrubstats[,c(4,6)]),type="upper")

#Our correlation matrix and corrplot indicate a medium negative corrlation shrub site and volume. Nothing to be concerned about.
```

#### Last bit of data wrangling
```{r Q1 - Data Wrangling Part 2}
#Ensure all variables are of appropriate class before running linear model
str(shrubstats)
#Convert site to a factor so that our model considers it to be categorical
shrubstats$site<-factor(shrubstats$site)

#Separate temperature and soil moisture so that we can run separate linear models (temperature as response for one, soil moisture as response for the other)

tempdata<-shrubstats[which(shrubstats$sensor=="temp"),]
moistdata<-shrubstats[which(shrubstats$sensor=="soil moisture"),] #sample size is much larger for soil moisture
```

### Model selection for temperature
```{r Q1 - Temp}
#Temperature

#Full linear model
tempfit1<-lm(tempdata$MeanMeasure~tempdata$treatment*tempdata$site*tempdata$microsite*tempdata$Volume)
summary(aov(tempfit1)) #only significant predictor is site
summary(tempfit1) #great r^2: 0.8772. But many not significant predictors.
tempfit2<-lm(tempdata$MeanMeasure~tempdata$site)
summary(aov(tempfit2)) #site remains significant.

#try removing site as we expect temperature to vary with latitude (which is involved with site)
tempfit3<-lm(tempdata$MeanMeasure~tempdata$treatment*tempdata$microsite*tempdata$Volume)
summary(aov(tempfit3)) #none significant.

boxplot(tempdata$MeanMeasure~tempdata$site)

#Check for outliers 
autoplot(tempfit1,c(1,2,3,6))
#There are cooks distance value over the standered threshold of 2. This points exhibits influence over the model and should not be included. Data should be transformed or a non-parametric test conducted. 
car::outlierTest(tempfit1)#Can use outlier test to double check
#Based on studentized residuals there is at least one outlier to be removed. 
```

#### Outlier shrub removed
```{r Q1 - Temp No Out}
#remove outlier from site 6 and run model
tempdatanoout<-tempdata[-which(tempdata$site==6),]
tempdatanooutfit1<-lm(tempdatanoout$MeanMeasure~tempdatanoout$treatment*tempdatanoout$site*tempdatanoout$microsite*tempdatanoout$Volume)
summary(aov(tempdatanooutfit1)) #none significant
summary(tempdatanooutfit1) #confirms none significant, however we do still see an effect of increasing temparature across sites, with site 5 having highest mean temperature, followed by site 2 and then site 1.

#Get F crit for site and volume for figures
#numerator degrees of freedom for main effect is #of groups in predictor-1, for interaction effect is (#of groups in predictor x -1)(#of groups in predictor y-1)(#of groups in predictor z-1)... denominator degrees of freedom is #points-(#of groups in predictor x)(#of groups in predictor y)(#of groups in predictor z)...-1
#e.g. here numerator degrees of freedom for site is 3-1=2, for volume is 1 (numerical predictor), and for denominator is 15-(2)(3)(2)(1)-1 (where there are 2 groups in treatment, 3 in site, 2 in microsite and we have our numerical predictor volume) = 15-12-1=2
qf(0.95,2,2) #site F-crit
qf(0.95,1,2) #volume F-crit

#Visualize response of temperature to site
boxplot(tempdatanoout$MeanMeasure~tempdatanoout$site)

#Check outliers and assumptions
autoplot(tempdatanooutfit1,c(1,2,3,6))
#There are still cooks distance value over the standard threshold of 2. These points exhibits influence over the model and ideally should not be included. However, due to our already small sample sizes we leave this model as the best one.
car::outlierTest(tempdatanooutfit1) #Can use outlier test to double check

```

#### Summary of temperature as response 
```{r Q1 - Temp Summary}
#How is latitude related to site?
cuyama$lat[which(cuyama$site==1)][1]
cuyama$lat[which(cuyama$site==2)][1]
cuyama$lat[which(cuyama$site==3)][1]
cuyama$lat[which(cuyama$site==4)][1]
cuyama$lat[which(cuyama$site==5)][1]
cuyama$lat[which(cuyama$site==6)][1] 
#Latitude increases with site. Thus, unexpectedly, temperature increases with latitude - but differences in latitude are small so this effect could be due to other climate factors.

#How is longitude related to site?
cuyama$long[which(cuyama$site==1)][1]
cuyama$long[which(cuyama$site==2)][1]
cuyama$long[which(cuyama$site==3)][1]
cuyama$long[which(cuyama$site==4)][1]
cuyama$long[which(cuyama$site==5)][1]
cuyama$long[which(cuyama$site==6)][1]  
#Looking at sites 1, 2 and 5, longitude decreases with site. Thus temperature increases with decreased longitude - as you move away from the coast.

#Conclusion: site has an effect on temperature but it is not statistically significant once outliers are removed.

#Look at the other predictors, even though they were all not statistically significant:

boxplot(tempdatanoout$MeanMeasure~tempdatanoout$microsite) #very similar temps between microsites, more variation in shrub microsite than open
boxplot(tempdatanoout$MeanMeasure~tempdatanoout$treatment) #temperatures higher for unclipped shrubs but again lots of variation so not statistically significant
plot(tempdatanoout$MeanMeasure~tempdatanoout$Volume) #bit of a trend of decreasing temperatures with increasing shrub volume, but too much variation/too small sample size to say for sure.

##OVERALL CONCLUSION: Temperature increases with site (increases with latitude and with decreasing longitude), decreases with increased volume, and is higher for unclipped shrubs. However, none of these trends are statistically significant when our outlier is removed (tempdatanooutfit1). When the outlier from site 6 is included in the model, the only statistically significant predictor of mean temperature is a main effect of site (tempfit1 and tempfit2).
```

### Model selection for soil moisture
```{r Q1 - Moist}
#Soil Moisture

#Remove negative soil moisture measures - they are not realistic since soil moisture is measured as a ratio of water volume to soil volume

moistdata<-moistdata[-which(moistdata$MeanMeasure<0),]

#Full linear model
moistfit1<-lm(moistdata$MeanMeasure~moistdata$treatment*moistdata$site*moistdata$microsite*moistdata$Volume)
summary(aov(moistfit1)) #all terms and interactions are significant
summary(moistfit1) #model has high r^2 - adjusted r^2 is 0.9994

#try removing site 3 (low sample size)
moistdatanos3<-moistdata[-which(moistdata$site==3),]

moistfit2<-lm(moistdatanos3$MeanMeasure~moistdatanos3$treatment*moistdatanos3$site*moistdatanos3$microsite*moistdatanos3$Volume)
summary(aov(moistfit2)) #all terms and interactions still significant
summary(moistfit2) #adjusted r^2 is 0.9998
#same results so include site 3 in our analysis

#check for outliers using autoplot

autoplot(moistfit1, c(1,2,3,6))#Definitely an outlier cook's distance is well over 10. There is something wrong there.
car::symbox(~MeanMeasure,data=moistdata)#Despite the failure to meet assumptions the symbox reccomends untransformed data
```

#### Outlier shrub removed
```{r Q1 - Moist No Out}
#try removing outlier (shrub 164)

moistdatanoout<-moistdata[-which(moistdata$shrub.ID==164),]

moistfit3<-lm(moistdatanoout$MeanMeasure~moistdatanoout$treatment*moistdatanoout$site*moistdatanoout$microsite*moistdatanoout$Volume)
summary(aov(moistfit3)) #now no terms are significant
summary(moistfit3) #r^2 went down a lot

#Get degrees of freedom and F crit for figures in poster
#site df: 6-1=5
#volume df: 1
#denominator df: 49-(2)(6)(2)(1)-1 = 49-24-1 = 24
qf(0.95,5,24)
qf(0.95,1,24)

```

#### Summary of soil moisture as response 
```{r Q1 - Moist Summary}

#full data - can't see trends because of extreme outlier.
boxplot(moistdata$MeanMeasure~moistdata$treatment)
boxplot(moistdata$MeanMeasure~moistdata$site)
boxplot(moistdata$MeanMeasure~moistdata$microsite)
plot(moistdata$MeanMeasure~moistdata$Volume)

#without outlier
boxplot(moistdatanoout$MeanMeasure~moistdatanoout$treatment) #unclipped has slightly lower soil moisture than clipped. Effect of predictor looks small.
boxplot(moistdatanoout$MeanMeasure~moistdatanoout$site) #site 5 has a lot of variation in soil moisture. Site 1 has highest median soil moisture. Effect of predictor looks small.
boxplot(moistdatanoout$MeanMeasure~moistdatanoout$microsite) #shrub has slightly lower soil moisture than open microsite. Effect of predictor looks small.
plot(moistdatanoout$MeanMeasure~moistdatanoout$Volume) #Effect of predictor looks small. Pretty constant soil moisture across shrub volumes.

##OVERALL CONCLUSION: When we do not remove possible outliers, all predictors and their interactions have a statistically significant effect on soil moisture and our model explains a lot of the variation in soil moisture content (high r^2). Once these possible outliers are removed, none of our predictors or their interactions have a statistically significant effect on soil moisture content. This latter result makes sense when we look at some basic plots of the data - the effects of predictors are overall small.
```

### Check assumptions for models in Q1
```{r Q1 - Model Assumptions}
#Check assumptions
#reference: https://data.library.virginia.edu/diagnostic-plots/

#Temperature model
plot(tempdatanooutfit1) #without outlier - given small sample size, linearity appears okay but have high variance in some residuals (residuals vs. fitted), normality is okay but sample size is very small so tough to tell how accurate this is (Normal Q-Q), homoscedasticity may not be well fit (scale-location), still appear to have outliers (Cook's distance)

#With the Shapiro-Wilkes test, our null hypothesis is that the residuals from our model are normally distributed - fit without outliers fails check (tempdatanooutfit1)
shapiro.test(residuals(tempdatanooutfit1))

#Moisture model
plot(moistfit3) #without outlier - linearity seems okay but some variation in residuals in the middle, fit to normality is again okay but not great (Normal Q-Q), homoscedasiticty assumption seems to be met but again some points with higher variation in residuals around the middle, still appear to have a couple of outliers present (residuals vs. leverage)

shapiro.test(residuals(moistfit3))
#both model fails Shapiro-Wilkes test - suggests model does not fit assumption of normality.

##An interesting next step for future analyses could be to run a generalized linear model with the data and see if results are similar/if fit to assumptions are improved.

```

# Question 2

## Does the effect of treatment on shrub volume vary geographically?

```{r Q2 - Data Wrangling}

#Filter shrubstats data - average over microsite - we want shrub volume as response, and site, treatment, temperature and soil moisture as predictors
shrubstatsq2<-summarize(group_by(shrubstats,shrub.ID,sensor,site,treatment,Volume), Mean=mean(MeanMeasure))

#separate measure into separate columns for temp and soil moisture
tempdat<-shrubstatsq2[c(2,6)]
tempdat[which(tempdat$sensor=="soil moisture"),]<-NA

moistdat<-shrubstatsq2[c(2,6)]
moistdat[which(moistdat$sensor=="temp"),]<-NA

#Put measurements back into dataset
shrubstatsq2<-cbind(shrubstatsq2[,-c(6)],temp=tempdat$Mean,moist=moistdat$Mean)

#remove negative soil moistures
shrubstatsq2<-shrubstatsq2[-which(shrubstatsq2$moist<0),]

#One row per shrub
shrubstatsq2[8,6]<-shrubstatsq2[9,6]
shrubstatsq2[11,6]<-shrubstatsq2[12,6]
shrubstatsq2[15,6]<-shrubstatsq2[16,6]
shrubstatsq2[22,6]<-shrubstatsq2[23,6]
shrubstatsq2[24,6]<-shrubstatsq2[25,6]
shrubstatsq2[29,6]<-shrubstatsq2[30,6]
shrubstatsq2[31,6]<-shrubstatsq2[32,6]
shrubstatsq2[34,6]<-shrubstatsq2[35,6]
shrubstatsq2<-shrubstatsq2[-c(9,12,16,23,25,30,32,35),]

#No missing data once filtered, do see some outliers for measures of shrub 164 by a quick look at the data.
```

#### Check for bias in sample sizes among predictors. Check for correlations among numerical predictors.
```{r Q2 - Predictors}

#Check for bias in sample sizes among type of sensor, site and treatment (predictors)

sum(complete.cases(shrubstatsq2$temp)) #We have a temperature measure for 11/31 total shrubs. 11 is a small sample size so could lead to bias.
sum(complete.cases(shrubstatsq2$moist)) #We have a soil moisture measure for 28/31 total shrubs.

sum(shrubstatsq2$treatment=="clipped") #sample size of clipped is 18/31 total shrubs, so we have a fairly even spread of clipped and unclipped shrubs

#Create a function for sample sizes for site
samplesizesiteq2func<-function(i){
  sum(shrubstatsq2$site==i)}
samplesizessiteq2<-sapply(1:6,samplesizesiteq2func) #sample sizes for site are even except for site 3 (only one data point). One data point is a very small sample size so we must keep this point in mind when completing our analyses.

#Check for correlations among numerical predictors (temperature, soil moisture).
t1<-as.data.frame(cor(shrubstatsq2[,c(6:7)],use ="complete.obs"))

#Use tableGrob from gridExtra package to create our table and store it as an object called "table"
table<-tableGrob(t1, rows = rownames(t1), cols = colnames(t1), theme = ttheme_default(rowhead=list(
        fg_params=list(fontface="bold"),
        bg_params=list(fill=c(NA,rep("grey80",4)))
    )))
grid.arrange(table)

corrplot(cor(shrubstatsq2[,c(6:7)],use ="complete.obs"),type="upper") #we see a high negative correlation between soil moisture and temperature

#Repeat above analysis without outlier shrub (shrub 164)
shrubstatsq2noout<-shrubstatsq2[-which(shrubstatsq2$shrub.ID==164),]

t2<-as.data.frame(cor(shrubstatsq2noout[,c(6:7)],use ="complete.obs"))

#Use tableGrob from gridExtra package to create our table and store it as an object called "table"
table<-tableGrob(t2, rows = rownames(t2), cols = colnames(t2), theme = ttheme_default(rowhead=list(
        fg_params=list(fontface="bold"),
        bg_params=list(fill=c(NA,rep("grey80",4)))
    )))
grid.arrange(table)

corrplot(cor(shrubstatsq2noout[,c(6:7)],use ="complete.obs"),type="upper")

#with the outlier removed, soil moisture and temperature have only a medium correlation.
```

### Model selection for volume as response
```{r Q2 - Initital Volume Models}

#Full model

volfit1<-lm(shrubstatsq2$Volume~shrubstatsq2$site*shrubstatsq2$treatment*shrubstatsq2$temp*shrubstatsq2$moist)
summary(aov(volfit1)) #R is not a fan, many observations deleted, p-values for terms not given. Try model without site. 

volfit2<-lm(shrubstatsq2$Volume~shrubstatsq2$treatment*shrubstatsq2$temp*shrubstatsq2$moist)
summary(aov(volfit2)) #R still deleted many observations. Likely due to the many NAs. Instead, we will try running our model separately for shrubs with both soil moisture and temperature measures, and for only soil moisture measures (only one plant has a temperature only measure so do not run a model for temperature only)

moistonlydata<-select(shrubstatsq2[which(is.na(shrubstatsq2$temp)==TRUE&is.na(shrubstatsq2$moist)==FALSE),],-temp)
moistandtempdata<-shrubstatsq2[which(is.na(shrubstatsq2$temp)==FALSE&is.na(shrubstatsq2$moist)==FALSE),]

#First, look at shrubs with only soil moisture measures:

volfitmoistonly1<-lm(moistonlydata$Volume~moistonlydata$site*moistonlydata$treatment*moistonlydata$moist)
summary(aov(volfitmoistonly1)) #significant predictors are site and interactions of site with treatment and site with moisture. Significant interaction means that we cannot trust our main effects results. So, we must keep the main effects for all of our predictors, as they are all involved in an interaction at some point.
summary(volfitmoistonly1) #high r^2 

#Get degrees of freedom and F crit for poster figures

#site df: 6-1=5
#treatment df: 2-1=1
#moisture df: 1 (numeric)
#site-temperature interaction df: 5(1)=5
#site-moisture interaction df: 5(1)=5
#denominator df: 20-(6)(2)(1)-1=7

qf(0.95,5,7) #applies to site and both significant interactions - since all numerator dfs are 5...not sure why test is showing different df values for interactions?

#keep only significant predictors in model (plus all main effects, as mentioned above):

volfitmoistonly2<-lm(moistonlydata$Volume~moistonlydata$site+moistonlydata$treatment+moistonlydata$moist+moistonlydata$site:moistonlydata$treatment+moistonlydata$site:moistonlydata$moist)
summary(aov(volfitmoistonly2)) #we now also have a significant main effect of moisture - but keep in mind that significant interaction means that we cannot trust our main effects results.
summary(volfitmoistonly2) #r^2 still good 

##Summary: Our best model for shrub volume as a response includes main effects of all predictors, and interactions between site and treatment and between site and soil moisture.

#Try running model without site 3(small sample size)

moistonlydatanos3<-moistonlydata[-which(moistonlydata$site==3),]

volfitmoistonlynos31<-lm(moistonlydatanos3$Volume~moistonlydatanos3$site*moistonlydatanos3$treatment*moistonlydatanos3$moist)
summary(aov(volfitmoistonlynos31)) #same terms significant as when we ran our model with the data point from site 3 included.

#Repeat for model with only significant predictors:
volfitmoistonlynos32<-lm(moistonlydatanos3$Volume~moistonlydatanos3$site+moistonlydatanos3$treatment+moistonlydatanos3$moist+moistonlydatanos3$site:moistonlydatanos3$treatment+moistonlydatanos3$site:moistonlydatanos3$moist)
summary(aov(volfitmoistonlynos32)) #Again same terms significant as when we ran our model with the data point from site 3 included.

##Summary: Use our model with the data point for site 3 unless removing this point drastically improves our model's fit to assumptions.

```

#### Visualizations of volume as response - only shrubs without temperature data included

```{r Q2 - Volume Moist Only Summary}
boxplot(moistonlydata$Volume~moistonlydata$site) #Generally, shrub volume decreases as site increases - thus, shrub volume decreases as latitude increases and longitude decreases
boxplot(moistonlydata$Volume~moistonlydata$treatment) #not much difference in shrub volume between clipped and unclipped treatments - suggests clipped shrubs do compensate for being clipped and end up as big as shrubs that were not clipped
plot(moistonlydata$Volume~moistonlydata$moist) #Shrub volume seems to increase as soil moisture content increases
```

#### Check final model assumptions
```{r - Moist Only Model Assumptions}
#Check model assumptions:

plot(volfitmoistonly2) #linearity seems okay but some variation in residuals in the middle (residuals vs fitted), fit to normality is pretty good (Normal Q-Q), homoscedasiticty assumption does not seem to have a good fit (scale-location), still appear to have an outlier present (residuals vs. leverage)

shapiro.test(residuals(volfitmoistonly2)) #meets assumption for normality (fail to reject null of Shapiro)

plot(volfitmoistonlynos32) #imporves linearity assumption a bit but otherwise model is about the same, assumptions wise, as our model with site 3.

##Conclusion: Use model with site 3.
```

#### Model selection for volume as response - data includes only shrubs that have both temperature and soil moisture measures.
```{r Q2 - Volume Moist and Temp Only}
#Now, look at shrubs with soil moisture and temperature measures:

#Small sample size - only 8 points

volfitmoistandtemp1<-lm(moistandtempdata$Volume~moistandtempdata$site*moistandtempdata$treatment*moistandtempdata$temp*moistandtempdata$moist) 
summary(aov(volfitmoistandtemp1)) #not a happy model - no residual degrees of freedom. Likely due to singularities between site and treatment.

#Try model without site:
volfitmoistandtemp2<-lm(moistandtempdata$Volume~moistandtempdata$treatment*moistandtempdata$temp*moistandtempdata$moist)
summary(aov(volfitmoistandtemp2)) #no significant predictors
summary(volfitmoistandtemp2) #r^2 is okay but adjusted isn't great

#Try model without treatment:
volfitmoistandtemp3<-lm(moistandtempdata$Volume~moistandtempdata$site*moistandtempdata$temp*moistandtempdata$moist) 
summary(aov(volfitmoistandtemp3)) #Again not a happy model - likely due to singularity in site (site 6) or small sample size
```

#### Outlier shrub removed
```{r Q2 - Volume Moist and Temp Only No Out}
#Now, try the model without our outlier (shrub 164, from site 6)

moistandtempdatanoout<-moistandtempdata[-which(moistandtempdata$shrub.ID==164),]

#Model without site:
volfitmoistandtempnoout1<-lm(moistandtempdatanoout$Volume~moistandtempdatanoout$treatment*moistandtempdatanoout$temp*moistandtempdatanoout$moist)
summary(aov(volfitmoistandtempnoout1)) #still no significant predictors
summary(volfitmoistandtempnoout1) #r^2 is okay but adjusted isn't great

#Try model without treatment:
volfitmoistandtempnoout2<-lm(moistandtempdatanoout$Volume~moistandtempdatanoout$site*moistandtempdatanoout$temp*moistandtempdatanoout$moist)
summary(aov(volfitmoistandtempnoout2)) #still not a happy model. Try combinations of predictors:

#Site and temperature
volfitmoistandtempnoout3<-lm(moistandtempdatanoout$Volume~moistandtempdatanoout$site*moistandtempdatanoout$temp)
summary(aov(volfitmoistandtempnoout3)) #site and temperature have significant main effects
summary(volfitmoistandtempnoout3) #r^2 is good

#Remove interaction (not significant):
volfitmoistandtempnoout4<-lm(moistandtempdatanoout$Volume~moistandtempdatanoout$site+moistandtempdatanoout$temp)
summary(aov(volfitmoistandtempnoout4)) #both predictors still significant
summary(volfitmoistandtempnoout4) #r^2 still good

#Site and soil moisture
volfitmoistandtempnoout5<-lm(moistandtempdatanoout$Volume~moistandtempdatanoout$site*moistandtempdatanoout$moist)
summary(aov(volfitmoistandtempnoout5)) #no significant predictors
summary(volfitmoistandtempnoout5) #r^2 not as good as previous model

#Tempearture and soil moisture
volfitmoistandtempnoout6<-lm(moistandtempdatanoout$Volume~moistandtempdatanoout$temp*moistandtempdatanoout$moist)
summary(aov(volfitmoistandtempnoout6))#no significant predictors
summary(volfitmoistandtempnoout6) #r^2 not good

##Conclusion: Use model with main effects of site and temperature -  best r^2 and adjusted r^2.
```

#### Summary of volume as response - only shrubs with both temperature and soil moisture data included
```{r Q2 - Volume Moist and Temp Only Summary}

##Summary: For shrubs with soil moisture and temperature measures, we must look at models with site and treatment as predictors separately. For the model with treatmment, soil moisture and temperature, none of our predictors show a significant effect on shrub volume. For our model with site,  soil moisture and temperature, we had to again look at effects of predictors separately. We found that the best model includes site and temperature as significant predcitors of shrub volume. However, it is important to keep in mind our small sample size that makes these results difficult to trust.

#Check basic visualizations:

boxplot(moistandtempdatanoout$Volume~moistandtempdatanoout$site) #shrub volume decreases as site increases - thus, shrub volume decreases as latitude increases and longitude decreases
boxplot(moistandtempdatanoout$Volume~moistandtempdatanoout$treatment) #similar shrub volume for clipped and unclipped treatments, but variation is larger for clipped
plot(moistandtempdatanoout$Volume~moistandtempdatanoout$moist) #small sample size but shrub volume seems to increase as soil moisture content increases
plot(moistandtempdatanoout$Volume~moistandtempdatanoout$temp) #small sample size but shrub volume seems to decrease as temperature increases
```

#### Check final model assumptions
```{r Q2 - Moist and Temp Only Model Assumptions}

plot(volfitmoistandtempnoout4) #assumptions are fairly well met - but sample size is small so it is hard to tell how much we can trust these results
shapiro.test(residuals(volfitmoistandtempnoout4)) #normality assumption met

```

# Summary: Best Models for Q1 and Q2
```{r Summary - Best Models}
##Q1

#Response: Temperature
#Possible predictors: Treatment, site, microsite, shrub volume (and their interactions)
#Remove outlier
#No significant predictors so best model is intercept only:
tempfit<-lm(tempdatanoout$MeanMeasure~1)

#Response: Soil moisture
#Possible predictors: Treatment, site, microsite, shrub volume (and their interactions)
#Remove outlier
#No significant predictors so best model is intercept only:
moistfit<-lm(moistdatanoout$MeanMeasure~1)

##Q2

#Response: Shrub Volume
#Possible predictors: Treatment, site, soil moisture (and their interactions)
#Significant predictors are site and interactions of site with treatment and site with moisture. Significant interaction means that we cannot trust our main effects results. So, we must keep the main effects for all of our predictors, as they are all involved in an interaction:

volfitmoistonly<-lm(moistonlydata$Volume~moistonlydata$site+moistonlydata$treatment+moistonlydata$moist+moistonlydata$site:moistonlydata$treatment+moistonlydata$site:moistonlydata$moist)


#Response: Shrub Volume
#Possible predictors: Treatment, site, soil moisture, temperature
#Very low sample size, could not run model with site and treatment together - had to separate
#For model with site we had to separate our 3 predictors into separate models
#Remove outlier shrub
#Site and temperature were significant predictors of shrub volume:

volfitmoistandtemp<-lm(moistandtempdatanoout$Volume~moistandtempdatanoout$site+moistandtempdatanoout$temp)

```
