---
title: "TimeSeries"
author: "Sam Gene"
date: "April 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(dplyr)
library(ggplot2)
library(ggfortify)
library(lubridate)
library(scales)
library(cowplot)
```

Load data, chack structure, and format - keep columns shrub ID, Volume, microsite, treatment, soil moisture, temperature, latitude and longitude, month, day, hour. Will also need to convert date formate and calculate volume.

```{r}
cuyama<-read.csv("Data/Cuyama.csv")
str(cuyama)
cuyama<-cuyama %>% mutate(Volume=(2/3)*pi*(x/2)*(y/2)*z,
                         Date=as.Date(paste(year, month, days,sep="-"), "%Y-%m-%d"),
                         Month=month.abb[month]) %>% select(shrub.ID,Volume,
                                                           lat,long,measure,
                                                           microsite,
                                                           treatment,sensor,
                                                           Date, hour,Month,site) 
```

Make soil moisture and temperature columns. Remove soil moisture negative values. These are either typos or monitor recording issues. Soil moisture should not be negative.
```{r}
TempDat<-cuyama %>% filter(sensor%in%'temp') %>% rename(Temp=measure) %>% filter(Temp>0)
MoistDat<-cuyama %>% filter(sensor%in%'soil moisture') %>% rename(Moisture=measure)%>% filter(Moisture > 0)
cuyama<-full_join(TempDat,MoistDat) %>% select(-sensor)
str(cuyama)

```
Create theme for plots
```{r}
CustomTheme<-function(){theme(axis.line = element_line(size = 0.6, linetype = "solid"),
                              axis.ticks = element_line(colour = "black",  size = 0.6), 
                              axis.title = element_text(size = 10),
                              axis.text = element_text(size = 10, colour = "black"),
                              axis.text.x = element_text(angle = 45,vjust = 0.6),
                              plot.title = element_text(hjust = 0.5),
                              panel.background = element_rect(fill = "white"),
                              legend.key=element_rect(fill="white"))}
```

Time series including all date and time stamps
```{r}
MonthMicrositeTemp <- ggplot(cuyama,aes(x=Date, y= Temp, color=microsite)) +
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +#Ploting individual points is too time consuming and messy even with adjusted alpha values. Use an estimated average line instead.
  xlab("Date") + ylab("Air Temperature (C)")+ggtitle("Microsite Comparison")+
  scale_colour_manual(name="Microsite types:",values=c("navy","forestgreen"))+
  scale_x_date(labels=date_format("%b %y"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))+
  CustomTheme()%+replace%theme(axis.title.x=element_blank())
#MonthMicrositeTemp
MonthMicrositeMoist <- ggplot(cuyama,aes(x=Date, y= Moisture, color=microsite)) +
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +
  xlab("Date") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("navy","forestgreen"))+
  scale_x_date(labels=date_format("%b %y"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))+
  CustomTheme()
#MonthMicrositeMoist
MonthTreatTemp <- ggplot(cuyama,aes(x=Date, y= Temp, color=treatment)) +
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +
  xlab("Date") + ylab("Air Temperature (C)")+ggtitle("Treatment Comparison")+
  scale_colour_manual(name=" Treatments:", values=c("steelblue","lightgreen"))+
  scale_x_date(labels=date_format("%b %y"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))+
  CustomTheme()%+replace%theme(axis.title.x=element_blank(),axis.title.y=element_blank())
#MonthTreatTemp
MonthTreatMoist <- ggplot(cuyama,aes(x=Date, y= Moisture, color=treatment)) +
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +
  xlab("Date") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  scale_x_date(labels=date_format("%b %y"))+
   guides(color=guide_legend(override.aes=list(fill=NA)))+
  CustomTheme()%+replace%theme(axis.title.y=element_blank())
#MonthTreatMoist
Comp<-plot_grid(MonthMicrositeTemp+theme(legend.position = "none"),
                MonthTreatTemp+theme(legend.position = "none"),
                MonthMicrositeMoist+theme(legend.position = "none"), 
                MonthTreatMoist+theme(legend.position = "none"),
                ncol=2,nrow=2, align=c("h"),labels=c("A","B","C","D"),hjust=-4)
legend_a<-get_legend(MonthMicrositeTemp +theme(legend.position="bottom"))
legend_b<-get_legend(MonthTreatTemp +theme(legend.position="bottom"))
plot_grid(Comp,legend_a,legend_b,ncol=1,rel_heights = c(1,0.1,0.05))

```
###Monthly averages for treatment 
#Did not use this version
```{r}
AveMonthTreat<-cuyama %>% group_by(Month,treatment) %>% summarize(TempMean=mean(na.omit(Temp)),
                                                                           MoistMean=mean(na.omit(Moisture)))
AveMonthTreat$Month<-factor(AveMonthTreat$Month,
                           levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
AveMonthTreatTemp <- ggplot(AveMonthTreat,aes(x=Month, y=TempMean, color=treatment)) +
  geom_point(na.rm=TRUE, size=2) +
  xlab("Month") + ylab("Air Temperature (C)")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  theme_classic()
AveMonthTreatTemp
AveMonthTreatMoisture <- ggplot(AveMonthTreat,aes(x=Month, y=MoistMean, color=treatment)) +
  geom_point(na.rm=TRUE, size=2) +
  xlab("Month") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  theme_classic()
AveMonthTreatMoisture
plot_grid(AveMonthTreatTemp, AveMonthTreatMoisture,ncol=1,nrow=2, align=c("h"))
```
##Monthly averages for microsite
#Did not use this version
```{r}
AveMonthMicro<-cuyama %>% group_by(Month,microsite) %>% summarize(TempMean=mean(na.omit(Temp)),
                                                                  MoistMean=mean(na.omit(Moisture)))
AveMonthMicro$Month<-factor(AveMonthMicro$Month,
                            levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
AveMonthMicroTemp <- ggplot(AveMonthMicro,aes(x=Month, y=TempMean, color=microsite)) +
  geom_point(na.rm=TRUE, size=2) +
  xlab("Month") + ylab("Air Temperature (C)")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  theme_classic()
AveMonthMicroTemp
AveMonthMicroMoisture <- ggplot(AveMonthMicro,aes(x=Month, y=MoistMean, color=microsite)) +
  geom_point(na.rm=TRUE, size=2) +
  xlab("Month") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  theme_classic()
AveMonthMicroMoisture
plot_grid(AveMonthMicroTemp, AveMonthMicroMoisture,ncol=1,nrow=2, align=c("h"))

```

##Some data wrangling to combine both treatment types into one column
```{r}
Treat <-cuyama %>% select(-microsite)
Micro<-cuyama %>% select(-treatment) %>% rename(treatment=microsite)
cuyama<-rbind(Treat,Micro)
cuyama$Month<-factor(cuyama$Month,
                           levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
```
##Combination of the two above composite plots
```{r}
AveMonth<-cuyama %>% group_by(Month,treatment) %>% summarize(n=n(),
                                                               TempMean=mean(na.omit(Temp)),
                                                               MoistMean=mean(na.omit(Moisture)),
                                                               Tempsd=sd(na.omit(Temp)),Moistsd=sd(na.omit(Moisture)),
                                                               seTemp=Tempsd/sqrt(n),seMoist=Moistsd/sqrt(n))

pd<-position_dodge(width=0.8)
AveMonthTemp <- ggplot(AveMonth,aes(x=Month, y=TempMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd)+
  geom_line(position=pd)+
  geom_errorbar(aes(ymin=TempMean-seTemp*qt(0.975,df=n-1),
                    ymax=TempMean+seTemp*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+
  xlab("Month") + ylab("Air Temperature (C)")+
  scale_colour_manual(name="Treatment and \n Microsite type",values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()%+replace%theme(axis.title.x=element_blank())

AveMonthMoisture<- ggplot(AveMonth,aes(x=Month, y=MoistMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd)+
  geom_line(position=pd)+
  geom_errorbar(aes(ymin=MoistMean-seMoist*qt(0.975,df=n-1),
                    ymax=MoistMean+seMoist*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+
  xlab("Month") + ylab("Soil Moisture")+
  scale_colour_manual(name="Treatment and \n Microsite type",values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()

Comp<-plot_grid(AveMonthTemp+theme(legend.position = "none"), 
                AveMonthMoisture+theme(legend.position = "none"), 
                align=c("h"),ncol=1,nrow=2,labels=c("A","B"),hjust=-1)

legend_c<-get_legend(AveMonthTemp +theme(legend.position="bottom"))
plot_grid(Comp,legend_c,ncol=,nrow=2,rel_heights = c(1,0.1))
```

```{r}
AveHourTreat<-cuyama %>% group_by(hour,treatment) %>% summarize(n=n(),TempMean=mean(na.omit(Temp)),
                                                                  MoistMean=mean(na.omit(Moisture)),
                                                               Tempsd=sd(na.omit(Temp)),Moistsd=sd(na.omit(Moisture)),
                                                               seTemp=Tempsd/sqrt(n),seMoist=Moistsd/sqrt(n))
AveHourTreat$hour<-as.factor(AveHourTreat$hour)
AveHourTreatTemp <- ggplot(AveHourTreat,aes(x=hour, y=TempMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd)+
  geom_line(position=pd)+
  geom_errorbar(aes(ymin=TempMean-seTemp*qt(0.975,df=n-1),
                    ymax=TempMean+seTemp*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+
  xlab("Time (hour)") + ylab("Air Temperature (C)")+
  scale_x_discrete(breaks=seq(0,23,2))+
  scale_colour_manual(values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()%+replace%theme(axis.title.x=element_blank(), axis.title.y=element_blank())
AveHourTreatMoisture <- ggplot(AveHourTreat,aes(x=hour, y=MoistMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd)+
  geom_errorbar(aes(ymin=MoistMean-seMoist*qt(0.975,df=n-1),
                    ymax=MoistMean+seMoist*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+
  geom_line(position=pd)+
  xlab("Time (hour)") + ylab("Soil Moisture")+
   scale_x_discrete(breaks=seq(0,23,2))+
  scale_colour_manual(values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()%+replace%theme(axis.title.y=element_blank())

Comp=plot_grid(AveHourTreatTemp+theme(legend.position = "none"), AveHourTreatMoisture+theme(legend.position = "none"),ncol=1,nrow=2, align=c("h"))

plot_grid(Comp,legend_c,ncol=1,nrow=2,rel_heights = c(1,0.1))
```

```{r}
##Alternatively what if we did a 4 pannel composite plot for month and hour?
Comp2<-plot_grid(AveMonthTemp+theme(legend.position = "none"), 
                AveHourTreatTemp+theme(legend.position = "none"), 
                AveMonthMoisture+theme(legend.position = "none"),
                AveHourTreatMoisture+theme(legend.position = "none"),
                align=c("h"),ncol=2,nrow=2,labels=c("A","B","C","D"),hjust=-4)
plot_grid(Comp2,legend_c,ncol=1,nrow=2,rel_heights = c(1,0.1))

```
##Heat map for fun
Because why not
```{r}
library(viridis)#Working with reds so need a colour blind friendly palette
library(ggExtra)
str(cuyama2)
max(cuyama2$Temp, na.rm=TRUE)
#cuyama2 <- cuyama %>% mutate(year = year(Date),
                  #month = month(Date, label=TRUE),
                  #day = day(Date)) %>% filter(treatment%in%'clipped')
cuyama2 <- cuyama %>%mutate(day = day(Date),Date=format(Date,format="%Y-%b")) %>% group_by(Date,day,hour,treatment) %>% summarize(Temp=mean(na.omit(Temp)))

treats<-unique(cuyama2$treatment)
p <-ggplot(cuyama2,aes(Date,hour,fill=Temp))+
  geom_tile() + 
  scale_fill_viridis(name="Hourly Air Temperature C",option ="C",discrete="FALSE")+ 
  facet_grid(day~treatment) + 
  scale_y_continuous(trans = "reverse", breaks = unique(cuyama2$hour))+
  theme_minimal(base_size = 8)+
  labs(title= paste("Hourly Temps"), x="Day", y="Hour")+
  CustomTheme()%+replace%theme(legend.position = "bottom",
                                 plot.title=element_text(size = 14),
                                 axis.text.y=element_text(size=6),
                                 strip.background = element_rect(colour="white"),
                                 axis.ticks=element_blank(),
                                 axis.text=element_text(size=7),
                                 legend.title=element_text(size=8),
                                 legend.text=element_text(size=6))+
  removeGrid()
p
cuyama3 <- cuyama %>% mutate(year = year(Date),
                  month = month(Date, label=TRUE),
                  day = day(Date))%>% group_by(year,month,day,hour,treatment) %>% summarize(Temp=mean(na.omit(Temp)))%>% filter(treatment%in%'clipped')

treats<-unique(cuyama3$treatment)

q<-ggplot(cuyama3,aes(day,hour,fill=Temp))+
  geom_tile(colour="white",size=0.1) + 
  scale_fill_viridis(name="Hourly Air Temperature C",option ="C",discrete="FALSE")+ 
  facet_grid(year~month) + 
  scale_x_continuous(breaks =c(1,10,20,31))+
  scale_y_continuous(trans = "reverse", breaks = unique(cuyama3$hour))+
  theme_minimal(base_size = 8)+
  labs(title= paste("Hourly Temps-",treats), x="Day", y="Hour")+
  CustomTheme()%+replace%theme(legend.position = "bottom",
                                 plot.title=element_text(size = 14),
                                 axis.text.y=element_text(size=6),
                                 strip.background = element_rect(colour="white"),
                                 axis.ticks=element_blank(),
                                 axis.text=element_text(size=7),
                                 legend.title=element_text(size=8),
                                 legend.text=element_text(size=6))+
  removeGrid()
q
#Last attempt
cuyama4 <- cuyama %>%mutate(day = day(Date),Date=format(Date,format="%y-%b")) %>% group_by(Date,day,hour,treatment) %>% summarize(Temp=mean(na.omit(Temp)))


z<-ggplot(cuyama4,aes(day,hour,fill=Temp))+
  geom_tile(colour="white",size=0.8) + 
  scale_fill_viridis(name="Hourly Air Temperature C",option ="C",discrete="FALSE")+ 
  facet_grid(treatment~Date) + 
  scale_x_continuous(breaks =c(1,10,20,31))+
  scale_y_continuous(trans = "reverse", breaks = unique(cuyama4$hour))+
  theme_minimal(base_size = 8)+
  labs(title= paste("Hourly Temps"), x="Day", y="Hour")+
  CustomTheme()%+replace%theme(legend.position = "bottom",
                                 plot.title=element_text(size = 14),
                                 axis.text.y=element_text(size=6),
                                 strip.background = element_rect(colour="white"),
                                 axis.ticks=element_blank(),
                                 axis.text=element_text(size=7),
                                 legend.title=element_text(size=8),
                                 legend.text=element_text(size=6))+
  removeGrid()
z

ggplot(cuyama2, aes(day, hour, fill= Temp)) + 
  geom_tile() +
  coord_equal()+
  scale_fill_viridis(discrete=FALSE)+
  CustomTheme()

```

##Timeseries with daily averages
#Not for use
```{r}
AveDateData<-cuyama %>% group_by(Date,treatment) %>%summarize(n=n(),TempMean=mean(na.omit(Temp)),
                                                    MoistMean=mean(na.omit(Moisture)),
                                                    Tempsd=sd(na.omit(Temp)),
                                                    Moistsd=sd(na.omit(Moisture)),
                                                    seTemp=Tempsd/sqrt(n),seMoist=Moistsd/sqrt(n))

AveDateTemp <- ggplot(AveDateData,aes(x=Date, y=TempMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd)+
  #geom_line(position=pd)+
  xlab("Date") + ylab("Air Temperature (C)")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  CustomTheme()

AveDateMoisture <- ggplot(AveDateData,aes(x=Date, y=MoistMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd)+
  #geom_line(position=pd)+
  xlab("Date") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  theme_classic()

Comp=plot_grid(AveDateTemp+theme(legend.position = "none"), AveDateMoisture+theme(legend.position = "none"),ncol=1,nrow=2, align=c("h"))
plot_grid(Comp,legend_c,ncol=1,nrow=2,rel_heights = c(1,0.1))
```


Play with later
```{r}
MeasureList=c("Temp","Moisture")
for (i in 1:2) {
  #Dat.i<-Dat4[Dat4$Variable==Param.i, ]
MeasureList[i]<-ggplot(cuyama,aes(x=Date, y=MeasureList[i], color=microsite)) +
  stat_smooth(na.rm=TRUE,se=TRUE,size=0.5)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +#Ploting individual points is too time consuming and messy even with adjusted alpha values. Use an estimated average line instead.
  xlab("Date") + ylab(expression("Air Temperature (C)")+
  scale_colour_manual(values=c("navy","forestgreen"))+
  scale_x_date(labels=date_format("%b %y"))+
  CustomTheme()
    ggplot(data=Dat.i) +
    geom_boxplot(aes(x=factor(as.factor(date)),y=Value,fill=factor(date))) + 
    scale_fill_manual(labels = c("2012", "2013"), values = c("lightgreen", "skyblue"))+
    theme(axis.ticks = element_line(colour = "black"),
          axis.title = element_text(size = 14),
          axis.text = element_text(size = 12, colour = "black"),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          axis.ticks.x=element_blank(),
          legend.text = element_text(size = 14),
          legend.title = element_text(size=16,colour="DarkRed"),
          panel.background = element_rect(fill = "white",colour="black"),
          legend.position = "right",
          plot.title=element_text(colour="firebrick",size=17))+
    labs( y = "Concentration (mg/L)",fill="Year") +ggtitle(Param.i)
  print(X2)
  dev.off()}
```