---
title: "TimeSeries"
author: "Sam Gene"
date: "April 14, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(dplyr)
library(ggplot2)
library(ggfortify)
library(lubridate)
library(scales)
library(cowplot)
```

Load data, chack structure, and format - keep columns shrub ID, Volume, microsite, treatment, soil moisture, temperature, latitude and longitude, month, day, hour. Will also need to convert date formate and calculate volume.

```{r}
cuyama<-read.csv("Data/Cuyama.csv")#Read in data
str(cuyama)#Checking the structure
cuyama<-cuyama %>% mutate(Volume=(2/3)*pi*(x/2)*(y/2)*z,#Calculate volume
                         Date=as.Date(paste(year, month, days,sep="-"), "%Y-%m-%d"),#re-format Date using lubridate
                         Month=month.abb[month]) %>% select(shrub.ID,Volume,#Select desired columns
                                                           lat,long,measure,
                                                           microsite,
                                                           treatment,sensor,
                                                           Date, hour,Month,site) 
```

Make soil moisture and temperature columns. Remove soil moisture negative values. These are either typos or monitor recording issues. Soil moisture should not be negative.
```{r}
TempDat<-cuyama %>% filter(sensor%in%'temp') %>% rename(Temp=measure) %>% filter(Temp>0)#seperate measure columns
MoistDat<-cuyama %>% filter(sensor%in%'soil moisture') %>% rename(Moisture=measure)%>% filter(Moisture > 0)
cuyama<-full_join(TempDat,MoistDat) %>% select(-sensor)#Join datasets back together
str(cuyama)#Check structure again

```

##Create a theme
Create a common theme for plots as a function so it can be added to all plots easily. I like theme_classic() so I based mine on that. 
```{r}
CustomTheme<-function(){theme(axis.line = element_line(size = 0.6, linetype = "solid"),#set axis line size and type
                              axis.ticks = element_line(colour = "black",  size = 0.6), #set tick colour and size
                              axis.title = element_text(size = 10),#axis title size
                              axis.text = element_text(size = 10, colour = "black"),#Text size and colour adjustment
                              axis.text.x = element_text(angle = 45,vjust = 0.6),#Adjust axis text angle and loacation
                              plot.title = element_text(hjust = 0.5),#Plot title height adjustment
                              panel.background = element_rect(fill = "white"),#Get rid of grey panneling
                              legend.key=element_rect(fill="white"))}##Get rid of legend key colour
```

#Time series including all dates and time stamps
### We didn't use this one but it is what helped us realize there were difference that might be more evident on a monthly or hourly basis. 
```{r}
MonthMicrositeTemp <- ggplot(cuyama,aes(x=Date, y= Temp, color=microsite)) + #Let's plot temp by date for microsite
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+#Estimating trend line 
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +#Ploting individual points is too time consuming and messy even with adjusted alpha values. Use an estimated average line instead.
  xlab("Date") + ylab("Air Temperature (C)")+ggtitle("Microsite Comparison")+
  scale_colour_manual(name="Microsite types:",values=c("navy","forestgreen"))+
  scale_x_date(labels=date_format("%b %y"))+#I want mont and tear as the date format. Individual days would be messy.
  guides(color=guide_legend(override.aes=list(fill=NA)))+#I really dislike the gray shading in the legend for the confidence intervals. I have removed it. 
  CustomTheme()%+replace%theme(axis.title.x=element_blank())#Remove redundant x-axis because this will be going in a composite plot
#MonthMicrositeTemp <- I checked the plot individually but hashtagged it out so you can just see it as a composite
MonthMicrositeMoist <- ggplot(cuyama,aes(x=Date, y= Moisture, color=microsite)) + #Plotting moisture by date in microsite
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +
  xlab("Date") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("navy","forestgreen"))+
  scale_x_date(labels=date_format("%b %y"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))+
  CustomTheme()
#MonthMicrositeMoist
MonthTreatTemp <- ggplot(cuyama,aes(x=Date, y= Temp, color=treatment)) +# Plot temp by date for treatmennts
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +
  xlab("Date") + ylab("Air Temperature (C)")+ggtitle("Treatment Comparison")+
  scale_colour_manual(name=" Treatments:", values=c("steelblue","lightgreen"))+ # I chose similar colours to the microsite type so that open and clipped correspond and shrub and unclipped correspons because they are essentially artificial versus natural versions of each other.
  scale_x_date(labels=date_format("%b %y"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))+
  CustomTheme()%+replace%theme(axis.title.x=element_blank(),axis.title.y=element_blank())# remove redundant labels. 
#MonthTreatTemp
MonthTreatMoist <- ggplot(cuyama,aes(x=Date, y= Moisture, color=treatment)) +#Plot moisture by date for treatments. 
  stat_smooth(na.rm=TRUE,se=TRUE,size=1)+
  #geom_point(na.rm=TRUE, size=1,alpha=I(0.2)) +
  xlab("Date") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("steelblue","lightgreen"))+
  scale_x_date(labels=date_format("%b %y"))+
   guides(color=guide_legend(override.aes=list(fill=NA)))+
  CustomTheme()%+replace%theme(axis.title.y=element_blank())
#MonthTreatMoist
Comp<-plot_grid(MonthMicrositeTemp+theme(legend.position = "none"),#Create a composite plot with no legends
                MonthTreatTemp+theme(legend.position = "none"),
                MonthMicrositeMoist+theme(legend.position = "none"), 
                MonthTreatMoist+theme(legend.position = "none"),
                ncol=2,nrow=2, align=c("h"),labels=c("A","B","C","D"),hjust=-4)
legend_a<-get_legend(MonthMicrositeTemp +theme(legend.position="bottom"))#Pull legend from a microsite plot
legend_b<-get_legend(MonthTreatTemp +theme(legend.position="bottom"))#Pull legend form a treatment plot
plot_grid(Comp,legend_a,legend_b,ncol=1,rel_heights = c(1,0.1,0.05))#Full compposite with legends incorporated. It's essentially a composite with a composite and some legends within it. 

```

##Some data wrangling to combine both treatment types into one column
The original versions of this graph had treatment and microsite seperate. Those versions are at the end of the markdown for curiosity. 
```{r}
Treat <-cuyama %>% select(-microsite)# remove microsite, name data Treat
Micro<-cuyama %>% select(-treatment) %>% rename(treatment=microsite)#remove treatment, name data treat, rename column treatemnt so that data can be rebound with the two as one column. 
cuyama<-rbind(Treat,Micro)#rebind datasets for new cuyama data. 
cuyama$Month<-factor(cuyama$Month,
                           levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
#Month is not being recognized as a proper date but I can reorder them here
```

##A composite plot combining the treatments and microsite types.
Values are averaged by month
```{r}
#Create means for moisture and temp with associated standard deviation, standard error, and n for 95% confidence intervals.
AveMonth<-cuyama %>% group_by(Month,treatment) %>% summarize(n=n(),
                                                               TempMean=mean(na.omit(Temp)),
                                                               MoistMean=mean(na.omit(Moisture)),
                                                               Tempsd=sd(na.omit(Temp)),Moistsd=sd(na.omit(Moisture)),
                                                               seTemp=Tempsd/sqrt(n),seMoist=Moistsd/sqrt(n))

pd<-position_dodge(width=0.8)#by creating pd, I can quickly move the points appart since some will overlap. 

AveMonthTemp <- ggplot(AveMonth,aes(x=Month, y=TempMean, color=treatment,group=treatment)) +#Plot for temperature
  geom_point(size=2,position=pd)+#Move the points apart for visibility
  geom_line(position=pd)+#make sure the lines go with them
  geom_errorbar(aes(ymin=TempMean-seTemp*qt(0.975,df=n-1),# add 95% confidence interval error bars using mutated columns
                    ymax=TempMean+seTemp*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+ #change colour, size and set position as pd
  xlab("Month") + ylab("Air Temperature (C)")+
  scale_colour_manual(name="Treatment and \n Microsite type",values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()%+replace%theme(axis.title.x=element_blank())#remove redundant axis title

AveMonthMoisture<- ggplot(AveMonth,aes(x=Month, y=MoistMean, color=treatment,group=treatment)) +#Plot fpr moisture
  geom_point(size=2,position=pd)+
  geom_line(position=pd)+
  geom_errorbar(aes(ymin=MoistMean-seMoist*qt(0.975,df=n-1),
                    ymax=MoistMean+seMoist*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+
  xlab("Month") + ylab("Soil Moisture (VWC)")+
  scale_colour_manual(name="Treatment and \n Microsite type",values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()

#Comp<-plot_grid(AveMonthTemp+theme(legend.position = "none"), 
                #AveMonthMoisture+theme(legend.position = "none"), 
                #align=c("h"),ncol=1,nrow=2,labels=c("A","B"),hjust=-1)

legend_c<-get_legend(AveMonthTemp +theme(legend.position="bottom"))#get legend for composite plot
#plot_grid(Comp,legend_c,ncol=,nrow=2,rel_heights = c(1,0.1))

#I did originally combine these two plots but they can be added to the following two for a 2X2 plot
```

Values averages by hour
```{r}
#Create means for moisture and temp with associated standard deviation, standard error, and n for 95% confidence intervals.
AveHourTreat<-cuyama %>% group_by(hour,treatment) %>% summarize(n=n(),TempMean=mean(na.omit(Temp)),
                                                                  MoistMean=mean(na.omit(Moisture)),
                                                               Tempsd=sd(na.omit(Temp)),Moistsd=sd(na.omit(Moisture)),
                                                               seTemp=Tempsd/sqrt(n),seMoist=Moistsd/sqrt(n))
AveHourTreat$hour<-as.factor(AveHourTreat$hour)#make hour a factor
AveHourTreatTemp <- ggplot(AveHourTreat,aes(x=hour, y=TempMean, color=treatment,group=treatment)) + #Hour temp plot
  geom_point(size=2,position=pd)+
  geom_line(position=pd)+
  geom_errorbar(aes(ymin=TempMean-seTemp*qt(0.975,df=n-1),# add CIs
                    ymax=TempMean+seTemp*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+
  xlab("Time (hour)") + ylab("Air Temperature (C)")+
  scale_x_discrete(breaks=seq(0,23,2))+
  scale_colour_manual(values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()%+replace%theme(axis.title.x=element_blank(), axis.title.y=element_blank())

AveHourTreatMoisture <- ggplot(AveHourTreat,aes(x=hour, y=MoistMean, color=treatment,group=treatment)) +#Soil moisture plot
  geom_point(size=2,position=pd)+
  geom_errorbar(aes(ymin=MoistMean-seMoist*qt(0.975,df=n-1),
                    ymax=MoistMean+seMoist*qt(0.975,df=n-1)),
                width=1,position=pd,colour="gray")+
  geom_line(position=pd)+
  xlab("Time (hour)") + ylab("Soil Moisture (VWC)")+
   scale_x_discrete(breaks=seq(0,23,2))+
  scale_colour_manual(values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()%+replace%theme(axis.title.y=element_blank())

#Comp=plot_grid(AveHourTreatTemp+theme(legend.position = "none"), AveHourTreatMoisture+theme(legend.position = "none"),ncol=1,nrow=2, align=c("h"))
#Combined two plots previously 

#plot_grid(Comp,legend_c,ncol=1,nrow=2,rel_heights = c(1,0.1))
```

#Poster plot
Alternativelyo my two 1X2 composite plots, I did a 4 pannel 2X2 composite plot for month and hour
```{r}
#combine plots and remove legends
Comp2<-plot_grid(AveMonthTemp+theme(legend.position = "none"), 
                AveHourTreatTemp+theme(legend.position = "none"), 
                AveMonthMoisture+theme(legend.position = "none"),
                AveHourTreatMoisture+theme(legend.position = "none"),
                align=c("h"),ncol=2,nrow=2,labels=c("A","B","C","D"),hjust=-4)
plot_grid(Comp2,legend_c,ncol=1,nrow=2,rel_heights = c(1,0.1))#Combine plots with legend extracted earlier. Alter heights so that legend takes up less room

```

#Unused figures and ideas

##Heat map for fun
Because why not
```{r}
library(viridis)#Working with reds so need a colour blind friendly palette
library(ggExtra)#Some extra ggplot help

#But we could totally do it as a loop 

TreatList<-c("clipped","unclipped","open","shrub")#Make a list
#MeasureList<-c("Temp", "Moisture")Takes way too long to run a double loop. Like over an hour. Instead lets try two plots
for (i in 1:length(TreatList)){
  Param.i<-TreatList[i]
  #Altering the data structure to make it more heatmap friendly.
  cuyama3 <- cuyama %>% mutate(year = year(Date),
                  month = month(Date, label=TRUE),
                  day = day(Date))%>% group_by(month,day,hour,treatment) %>% summarize(Temp=mean(na.omit(Temp)),Moisture=mean(na.omit(Moisture)))%>% filter(treatment%in%Param.i)

q<-ggplot(cuyama3,aes(day,hour,fill=Temp))+
  geom_tile(colour="white",size=0.1) + #To split up our individual tiles
  scale_fill_viridis(name="Air Temperature (C)",option ="C",discrete="FALSE")+ #Colour blind friendly scale
  facet_grid(.~month) + #use facets to make columns by month. I originally also had a secon x axis by year but it is less visually appealing so it is now averaged by month. 
  scale_x_continuous(breaks =c(1,10,20,31))+#Set scale breaks
  scale_y_continuous(trans = "reverse", breaks = unique(cuyama3$hour))+#Set breaks
  theme_minimal(base_size = 8)+
  labs(title= paste("Hourly Air Temperature-" ,Param.i), x="Day", y="Hour")+# Set labs
  CustomTheme()%+replace%theme(legend.position = "bottom",#Adjust the themes
                                 plot.title=element_text(size = 12),
                                 axis.text.y=element_text(size=6),
                                 strip.background = element_rect(colour="white"),
                                 axis.ticks=element_blank(),
                                 axis.text=element_text(size=7),
                                 legend.title=element_text(size=8),
                                 legend.text=element_text(size=6))+
  removeGrid()

p<-ggplot(cuyama3,aes(day,hour,fill=Moisture))+#Repeat for moisture double since loop took hours so wasn't very rerunable
  geom_tile(colour="white",size=0.1) + 
  scale_fill_viridis(name="Soil Moisture",option ="C",discrete="FALSE")+ 
  facet_grid(.~month) + 
  scale_x_continuous(breaks =c(1,10,20,31))+
  scale_y_continuous(trans = "reverse", breaks = unique(cuyama3$hour))+
  theme_minimal(base_size = 8)+
  labs(title= paste("Hourly Soil Moisture -" ,Param.i), x="Day", y="Hour")+
  CustomTheme()%+replace%theme(legend.position = "bottom",
                                 plot.title=element_text(size = 12),
                                 axis.text.y=element_text(size=6),
                                 strip.background = element_rect(colour="white"),
                                 axis.ticks=element_blank(),
                                 axis.text=element_text(size=7),
                                 legend.title=element_text(size=8),
                                 legend.text=element_text(size=6))+
  removeGrid()
print(p)#Print soil moisture plots
print(q)#Print temp plots
}
```
##Timeseries with daily averages
####Not for use
An alternate version of the first plot by date where measures were averaged by day
```{r}
AveDateData<-cuyama %>% group_by(Date,treatment) %>%summarize(n=n(),TempMean=mean(na.omit(Temp)),
                                                    MoistMean=mean(na.omit(Moisture)),
                                                    Tempsd=sd(na.omit(Temp)),
                                                    Moistsd=sd(na.omit(Moisture)),
                                                    seTemp=Tempsd/sqrt(n),seMoist=Moistsd/sqrt(n))

AveDateTemp <- ggplot(AveDateData,aes(x=Date, y=TempMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd,alpha=0.2)+
  #geom_line(position=pd)+
  xlab("Date") + ylab("Air Temperature (C)")+
  scale_colour_manual(values=c("steelblue","lightgreen","navy","forestgreen"))+
  CustomTheme()

AveDateMoisture <- ggplot(AveDateData,aes(x=Date, y=MoistMean, color=treatment,group=treatment)) +
  geom_point(size=2,position=pd,alpha=0.2)+
  #geom_line(position=pd)+
  xlab("Date") + ylab("Soil Moisture")+
  scale_colour_manual(values=c("steelblue","lightgreen","navy","forestgreen"))+
  theme_classic()

Comp=plot_grid(AveDateTemp+theme(legend.position = "none"), AveDateMoisture+theme(legend.position = "none"),ncol=1,nrow=2, align=c("h"))
plot_grid(Comp,legend_c,ncol=1,nrow=2,rel_heights = c(1,0.1))
```

##By site for curiosity
####Not for use
```{r}
SiteDateData<-cuyama %>% group_by(Date,site) %>%summarize(TempMean=mean(na.omit(Temp)),
                                                                        MoistMean=mean(na.omit(Moisture)))
SiteDateTemp <- ggplot(SiteDateData,aes(x=Date, y=TempMean, color=as.factor(site))) +#make site a factor and set colour to site
  geom_point(na.rm=TRUE, size=2,alpha=0.5) +
  xlab("Date") + ylab("Air Temperature (C)")+
  scale_colour_viridis(discrete = T,option="D")+#Use colour blind friendly scale 
  CustomTheme()
SiteDateTemp
SiteDateMoist <- ggplot(SiteDateData,aes(x=Date, y=MoistMean, color=as.factor(site))) +#make site a factor and set colour to site
  geom_point(na.rm=TRUE, size=2,alpha=0.5) +
  xlab("Date") + ylab("Soil Moisture")+
  scale_colour_viridis(discrete = T,option="D")+#Use colour blind friendly scale 
  CustomTheme()
SiteDateMoist
#Something different with site 6 to cause that spike. But too many points spike for it to be a typo or minor error. 
```

